Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-05 paper/disc20clustering.tex
Missing parts of LB

\begin{theorem}
	\label{th:lowerbound}
	The competitive ratio of any deterministic online algorithm for \OBRP{} is  $\Omega(k\cdot \ell)$ for any $k\geq 3$ and for any $\ell \geq 2$.
\end{theorem}

@@

	We distinguish a cluster $S$, where we form a ground set of size $k-1$.
	We never dismiss any constructed ground set.
	Each cluster host exactly $k$ nodes, hence in any configuration of \ALG, there exists a single alone node in $S$.

	We construct a ground set containing this alone node (and some other node).
	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S$.
	We construct a ground set $\{x_0, y_0\}$.
	\ALG has them split, and the adversary produces external requests until these are collocated.
	$S_1$ already contains a ground set of size $k-1$ (that it cannot split), hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.

	We continue to construct ground sets containing the current alone node from $S$.
	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
	\ALG must replace $x_0$ with another single node (that we name $x_1$).
	Then, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t. and $C \neq \{x_0,y_0\}$~$I(C) = I(x_1)$.
	Continuing this, we observe a sequence $X := x_0, x_1, x_2\dots$ of alone nodes that visit $S$.
	(Note that subsequent ground sets are different than the first ground set $\{x_0, y_0\}$, whose nodes were initially in different clusters.)

	We claim that this procedure (of constructing ground sets) can continue as long as a perfect partition of ground sets still exists.
	%and we construct a sequence of ground sets $X := \langle \{x_0, y_0\}, \{x_1, y_2\}, \{x_2, y_2\}, \ldots \rangle$ of length $\Omega(k \cdot \ell)$.
	Consider the moment of creating a ground set $\{x_i, y_i\}$.
	If a perfect partition of ground sets still exists, then
	\ALG{} must collocate $\{x_i,y_i\}$ in a~cluster other than $S$.
	To this end, it replaces $x_i$ with another alone node $x_{i+1}$.
	To continue, we must guarantee that
	\maciek{elaborate}

	We end this procedure prematurely (before all nodes visit $S$) to keep \OPT's cost low.
	The sequence must end when no perfect partition of ground sets exists ---
	however, we end it when \emph{two more} ground sets are still possible to give.
	Precisely, we repeat constructing ground sets,
	until there are only two alone nodes left that originate from the same cluster.
	We denote these two nodes by $x^*, y^*$.
	Observe that at this point there are at most $\ell+1$ alone nodes left:
	otherwise, by pigeon hole principle, there would be more pairs of alone nodes that were initially in the same cluster.

	\ALG incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
	We lower bound the length of the sequence $X := x_0, x_1,\dots$.
	\maciek{TODO}
	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate on exclude. Just sum up the real cost.}.

@@

	The competitive ratio is then: $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
\end{proof}

>>>>

\begin{theorem}
	\label{th:lowerbound}
	The competitive ratio of any deterministic online algorithm for \OBRP{} is  $(k\cdot \ell - k - \ell)/2$ for any $k\geq 3$ and for any $\ell \geq 2$.
\end{theorem}

@@

	First, where we form a ground set of size $k-1$ on an arbitrarily chosen cluster $S$.
	%We never dismiss any constructed ground set.
	Each cluster host exactly $k$ nodes, hence in any configuration of \ALG, there exists a single alone node in $S$.

	Then, we construct a ground set containing this alone node (and some other node).
	Precisely,
	we construct a ground set $\{x_0, y_0\}$, 
	where $x_0$ is the alone node from $S$ and $y_0$ is an arbitrary node that does not reside in $S$.
	\ALG has them split, and the adversary produces external requests until these are collocated.
	$S$ already contains a ground set of size $k-1$ (that it cannot split), hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.

	We continue to construct ground sets containing the node that is currently alone in $S$.
	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
	\ALG must replace $x_0$ with another single node (that we name $x_1$).
	Consequently, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x_1)$.
	Repeating this, we observe a sequence $X := x_0, x_1, x_2\dots$ of alone nodes that visit $S$.
	(Note that subsequent ground sets are different than the first ground set $\{x_0, y_0\}$, whose nodes were initially in different clusters.)

	We claim that constructing such ground sets can continue as long as at least $\ell$ alone node exists.
	Fix the moment just before the creation of a ground set with the node $x_i$.
	We show that if at least $\ell$ alone nodes exists (including $x_i, y_i$), then there exists a perfect partition of ground sets.

	To show this claim, we consider a configuration, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
	In this configuration, each cluster contains:  at most one ground set larger than $1$, possibly some alone nodes, and possibly the ground set $\{x_0, y_0\}$.
	If $I(x_i)$ contains an alone node, then we simply obtain a perfect partition by swapping it with $x_i$.
	Otherwise, $\{x_0, y_0\}$ is present on $I(x_i)$ (a single ground set cannot occupy $I(x_i)$, as without $x_i$ its size is at most $k-1$).
	In this case, we first swap $\{x_0, y_0\}$ with two alone nodes that occupy the same cluster.
	To see that such two nodes must exist, observe that we have at most $\ell-1$ clusters with alone nodes, and by the assumption we have $\ell$ alone nodes.
	After this swap, we migrate $x_i$ to $I(x_i)$, evicting a lone node to $S$, and we conclude that the perfect partition exists.

	To keep \OPT's cost low, we end constructing ground sets prematurely --- while there are still $\ell+1$ alone nodes (i.e., not performing two last constructions).
	Then, there two alone nodes exists that originate from the same cluster (we have $\ell+1$ alone nodes and $\ell$ clusters).
	We denote these two nodes by $x^*, y^*$.

@@

	\ALG performs at least one swap for each ground set $\langle \{x_0, y_0\}, \{x_1, y_1\}, \{x_2, y_2\}, \ldots \rangle$.
	We start with $k \cdot \ell - (k-1)$ single nodes (a single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with $\ell+1$ nodes.
	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - l) / 2$.
\end{proof}
